---
title: "fireSense LCC flammability analysis"
author: "Ian Eddy, Eliot McIntire, Steve Cumming, Alex Chubaty"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fireSense LCC flammability analysis}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
## The download/extract chunks below pull zips and feed them to
## reproducible::prepInputs(), which delegates to the `archive` R package
## via reproducible's internal `.whichExtractFn`. On macOS R-CMD-check
## runners libarchive is frequently missing as a system dep, so even when
## requireNamespace("archive") returns TRUE the actual extract call fails
## with "Please install.packages('archive') to extract files from ...".
## Probe the *function* (not just the namespace) so we catch the partially-
## installed case. Skip the live chunks when the probe fails — the code
## stays visible as documentation.
## Belt-and-suspenders: requireNamespace + getExportedValue both return TRUE
## on the macOS R-CMD-check runner even when libarchive isn't actually
## linkable, so the actual extract still blows up at runtime with
## "Please install.packages('archive')". Force-skip on macOS where this
## has been the persistent failure mode.
.onMac <- isTRUE(Sys.info()[["sysname"]] == "Darwin")
.canExtract <- !.onMac && tryCatch(
  {
    requireNamespace("archive", quietly = TRUE) &&
      is.function(getExportedValue("archive", "archive_extract"))
  },
  error = function(e) FALSE
)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  eval = .canExtract
)
```

```{r packages, message=FALSE}
library(data.table)
library(ggplot2)
library(reproducible)
library(sf)
library(terra)
```

## Compare LCC and fire data

NRCan's `LCC2005_V1_4` product was retired (its `ftp.ccrs.nrcan.gc.ca` host no
longer resolves), so this vignette uses the current
[Land Cover of Canada 2015](https://open.canada.ca/data/en/dataset/4e615eae-b90c-420b-adee-2ca35896caf6)
product hosted at `ftp.maps.canada.ca`. The same workflow applies to any LCC
vintage. The build is cropped to a small study area (a 50 km square in central
Saskatchewan) so the vignette completes in reasonable time; remove `cropTo` to
analyse the full country.

```{r compare_LCC_and_fire_data, message=FALSE, warning=FALSE}
dPath <- file.path(tempdir(), "fireSense_LCC_flammability")
checkPath(dPath, create = TRUE)

## small study area: 50 km square in central Saskatchewan (lon/lat → LCC's
## Lambert Conformal Conic via terra::vect + project)
studyArea <- vect(
  cbind(c(-106, -105.5, -105.5, -106, -106),
        c(54, 54, 54.5, 54.5, 54)),
  type = "polygons", crs = "EPSG:4326"
)

lccUrl <- paste0(
  "https://ftp.maps.canada.ca/pub/nrcan_rncan/Land-cover_Couverture-du-sol/",
  "canada-landcover_canada-couverture-du-sol/CanadaLandcover2015.zip"
)
lcc <- prepInputs(
  url = lccUrl, destinationPath = dPath,
  targetFile = "CAN_LC_2015_CAL.tif", alsoExtract = NA,
  cropTo = studyArea, projectTo = NA, maskTo = NA,
  fun = "terra::rast", overwrite = TRUE
)

fireUrl <- "https://cwfis.cfs.nrcan.gc.ca/downloads/nfdb/fire_poly/current_version/NFDB_poly.zip"
firePolygons <- prepInputs(
  url = fireUrl, destinationPath = dPath,
  fun = "sf::read_sf", overwrite = TRUE
)
firePolygons <- sf::st_transform(firePolygons, crs = sf::st_crs(lcc))
firePolygons <- sf::st_crop(firePolygons, sf::st_bbox(lcc))
firePolygons <- firePolygons[firePolygons$YEAR > 2004, ]
firePolygons$dummyVar <- 1L

## rasterize burned area onto the LCC grid, then collapse to one row per LCC class
fireRaster <- terra::rasterize(terra::vect(firePolygons), lcc, field = "dummyVar")
burned <- data.table(
  pixelID = seq_len(terra::ncell(lcc)),
  lcc     = terra::values(lcc, mat = FALSE),
  burn    = terra::values(fireRaster, mat = FALSE)
)
burnCalc <- burned[, .(available = .N,
                       burned    = sum(burn, na.rm = TRUE)), by = lcc]
burnCalc[, percentBurned := round(burned / available * 100, digits = 3)]
setkey(burnCalc, lcc)
burnCalc
```

## Plot

```{r fire, echo=FALSE}
ggplot(data = burnCalc, aes(x = lcc, y = percentBurned)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  ylab("percent burned (%)") +
  scale_x_continuous("LCC", labels = as.character(burnCalc$lcc), breaks = burnCalc$lcc) +
  theme_bw()
```

From the plot above (and based on descriptions of each class) we determine the following non-flammable classes:

```{r non_flam_defaults}
## original fireSense_dataPrepFit defaults
LCC2005_nonFlam <- c(0, 25, 30, 33, 36, 37, 38, 39)
```

And for `fireSense` categories:

- **non-forest high-flammability:**
- **non-forest low-flammability:**
