---
title: "fireSense tutorial"
author: "Jean Marchal, Eliot McIntire, Tati Micheletti, Alex Chubaty"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
references:
- id: Marchal2017
  type: article-journal
  author:
  - family: Marchal
    given: Jean
  - family: Cumming
    given: Steve G.
  - family: McIntire
    given: Eliot J. B.
  issued:
  - year: '2017'
    month: '1'
  title: '[Exploiting Poisson additivity to predict fire frequency from maps of
    fire weather and land cover in boreal forests of Québec, Canada]{.nocase}'
  container-title: Ecography
  page: '200-209'
  volume: '40'
  issue: '1'
  abstract: Predictive models of fire frequency conditional on weather and land cover
    are essential to assess how future cover-type distributions and weather conditions
    may influence fire regimes. We modelled the effects of bottom-up variables (e.g.
    land cover) and top-down variables (e.g. fire weather) simultaneously with data
    aggregated or interpolated to spatial and temporal units of 100 km2 and 1 yr in
    the boreal forest of Québec, Canada. For models of human-caused fires, we used
    road density as a surrogate for human access and behaviour. We exploited the additive
    property of Poisson distributions to estimate cover-type specific fire count rates,
    which would normally not be possible with data of this spatial resolution. We
    used piecewise linear functions to model nonlinear relations between fire weather
    and fire frequency for each cover-type simultaneously. The estimated conditional
    rates may be considered as expected mean counts per unit area and time. It follows
    that these rates can be rescaled to arbitrary spatial and temporal extents. Our
    results showed fire frequency increased nonlinearly as aridity increased and more
    quickly in disturbed areas than other types. Road density exerted the strongest
    influence on the frequency of human-caused fires, which were positively correlated
    with road density. The estimates may be used to parameterize the fire ignition
    component of spatial simulation models, which often have a resolution different
    from that at which the data were collected. This is an essential step in incorporating
    biotic and abiotic feedbacks, land-cover dynamics, and climate projections into
    ecological forecasting. The insight into the power of Poisson additivity to reveal
    high-resolution ecological processes from low-resolution data could have applications
    in other areas of ecology.
  DOI: 10.1111/ecog.01849
  ISSN: '09067590'
- id: Marchal2019
  type: article-journal
  author:
  - family: Marchal
    given: Jean
  - family: Cumming
    given: Steven G.
  - family: McIntire
    given: Eliot J. B.
  issued:
  - year: '2019'
    month: '5'
  title: '[Turning Down the Heat: Vegetation Feedbacks Limit Fire Regime Responses
    to Global Warming]{.nocase}'
  container-title: Ecosystems
  URL: http://link.springer.com/10.1007/s10021-019-00398-2
  DOI: 10.1007/s10021-019-00398-2
  ISSN: '1432-9840'
- id: Schoenberg2003
  type: article-journal
  author:
  - family: Schoenberg
    given: Frederic Paik
  - family: Peng
    given: Roger
  - family: Woods
    given: James
  issued:
  - year: '2003'
    month: '9'
  title: '[On the distribution of wildfire sizes]{.nocase}'
  container-title: Environmetrics
  page: '583-592'
  volume: '14'
  issue: '6'
  abstract: tapered Pareto
  keyword: Pareto,Tapered Pareto,Wildfires,pareto,tapered pareto,wildfires
  URL: http://doi.wiley.com/10.1002/env.605
  DOI: 10.1002/env.605
  ISSN: '1180-4009'
vignette: >
  %\VignetteIndexEntry{fireSense tutorial}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  eval = FALSE ## TODO: temporary until I can rework this vignette to actually run
)
```

```{r init}
library(Require)
Require(c("magrittr", "data.table", "dplyr", "pemisc", "PtProcess", "quickPlot",
          "raster", "sf", "sp", "SpaDES.core", "SpaDES.tools", "spatstat"))

options(reproducible.useCache = FALSE) # Turn off caching

moduleDir <- checkPath(file.path(tempdir(), "modules"), create = TRUE)
outputDir <- checkPath(file.path(tempdir(), "outputs"), create = TRUE)

setPaths(
  modulePath = moduleDir,
  outputPath = outputDir
)

## TODO: download modules to moduleDir -- use git or downloadModule() ???

normalizeRaster <- function(x) { ## TODO: use pemisc version (#9)
  # Normalize raster values
  (x - minValue(x)) / (maxValue(x) - minValue(x))
}
```

# Introduction

`fireSense` is a spatially explicit landscape fire model, that is, it simulates the fire disturbance in the landscape.
`fireSense` is implemented as a group of `SpaDES` modules.
A main asset of `fireSense` is that it can be parameterized directly from data using systematic methods while incorporating sensitivity to different environmental drivers.

To better represent how environmental drivers control fire, `fireSense` breaks down the fire process into three stages: ignition, escape, and spread.
Sensitivity to environmental drivers can be introduced at each of these stages and these factors need not necessarily be identical for all stages.
This way each module can be adapted to your own problem and used to answer your own questions.

In order to facilitate the parameterization of `fireSense` from data using systematic methods we have created SpaDES modules dedicated to these tasks:

- `fireSense_IgnitionFit`
- `fireSense_EscapeFit`
- `fireSense_SpreadFit`

These follow the way `fireSense` breaks down the fire process and are named accordingly. For example, `fireSense_IgnitionFit` implements methods to fit the statistical model of fire ignitions.
The `fireSense_IgnitionFit` module implements methods described in Marchal et al. [-@Marchal2017], `fireSense_EscapeFit` and `fireSense_SpreadFit` implement methods described in Marchal et al. [-@Marchal2019].
Each of these modules is associated with an additional module to facilitate the production of predictive maps (rasters) or data sets (data.frames).
To continue with the example of fire ignitions, the module `fireSense_IgnitionPredict` is associated with the `fireSense_IgnitionFit` module.
One possible use of the `fireSense_IgnitionPredict` module would be to generate rasters describing in a spatially explicit way the predicted ignition rates.

In addition of these modules, we implemented the `fireSense_SizeFit` module to fit statistical model of the fire size distribution using the tapered Pareto distribution [@Schoenberg2003].
The use of `fireSense_SizeFit` is not a requirement because these outputs are not required by `fireSense`.
Although the use of `fireSense_SizeFit` is optional, it can be used to reduce the dimensionality of the problem, which helps speed up and facilitate convergence when using the . module later.
Later in this tutorial we will show how the outputs of the `fireSense_SizeFit` module can be used as inputs of the `fireSense_SpreadFit` module.

Alternatively, the parameter of the tapered Pareto distribution can be used as inputs of the spread model assuming these capture the effect of environmental drivers of fire sizes.

We start by fitting the spread model using the `fireSense_SpreadFit` module.
In order to fit this model we need data about the shape of the fire size distribution.
We can predict it using the `fireSense_SizePredict` module and fit the model to our study region using the `fireSense_SizeFit` module.

This tutorial is intended to provide examples to run all `fireSense` modules without the need to download actual data.
Depending on the fact that some of them have dependencies.

```{r create_data, message=FALSE}
# Create a raster template describing the study area
raster_width <- raster_height <- 100
r_template <- raster(matrix(nrow = raster_height, ncol = raster_width))

# Create rasters describing environmental drivers controlling fire
# Here we use for example the land cover and the weather

set.seed(123)

## Landcover -- these should be "proportion of that single cover type", e.g., Proportion Deciduous
landTypeOne <- gaussMap(r_template, scale = 300, var = 300)
landTypeOne <- normalizeRaster(landTypeOne)
landTypeTwo <- 1 - landTypeOne

## Weather
weather <- gaussMap(r_template, scale = 300, var = 3000)

# Create low resolution rasters of the environmental drivers
# These are used for fitting the ignition model. The main advantages
# of fire modeling using low resolution are described in Marchal et al. 2017. 
# In short, unlike high-resolution modeling, fire locations need not be known
# precisely and do not require data at scales where local homogeneity of 
# landcover can be assumed.

landTypeOneLowRes <- aggregate(landTypeOne, fact = 25, fun = mean)
landTypeTwoLowRes <- 1 - landTypeOneLowRes

weatherLowRes <- aggregate(weather, fact = 25, fun = mean)

# Let's use a known pattern / equation
nbFires <- 500
fireLocations <- rpoint(nbFires, f = function(x, y, ...) 300 * y) # Introduce some spatial variability: more fires in the North
fireLocations <- as(fireLocations, "SpatialPoints")

#landcoverDataLowRes <- extract(landTypeOneLowRes, fireLocations, cellnumbers = TRUE) %>%
#  as_tibble %>%
#  rename(cell_id = 1, landtype_1_pp = 2) %>% # Rename columns using column index
#  distinct(cell_id, landtype_1_pp) %>%
#  mutate(landtype_2_pp = 1 - landtype_1_pp)  # Add a column with the 2nd landcover type

#weatherDataLowRes <- extract(weatherLowRes, fireLocations, cellnumbers = TRUE) %>%
#  as_tibble %>% 
#  rename(cell_id = 1, weather = 2) %>% # Rename columns using column index
#  distinct(cell_id, weather)

# Create the dataset necessary to fit the statistical model of fire ignitions
#dataFireSense_IgnitionFit <- landcoverDataLowRes %>%
#  group_by(cell_id) %>%
#  summarise(n_fires = n()) %>%
#  right_join(landcoverDataLowRes) %>%
#  left_join(weatherDataLowRes) %>%
#  mutate(n_fires = ifelse(is.na(n_fires), 0, n_fires))

# Converting the above from tibble to data.table 
dataFireSense_IgnitionFit <- data.table(
  extract(landTypeOneLowRes, fireLocations, cellnumbers = TRUE),
  landtype_2_pp = extract(landTypeTwoLowRes, fireLocations, cellnumbers = FALSE),
  weather = extract(weatherLowRes, fireLocations, cellnumbers = FALSE)
)
setnames(dataFireSense_IgnitionFit, old = "layer", new = "landtype_1_pp")
dataFireSense_IgnitionFit <- dataFireSense_IgnitionFit[, list(
  n_fires = .N, landtype_1_pp = landtype_1_pp[1], 
  landtype_2_pp = landtype_2_pp[1], weather = weather[1]
), by = cells]

dataFireSense_EscapeFit <- data.table(
  landtype_1_pp = extract(landTypeOne, fireLocations, cellnumbers = FALSE),
  landtype_2_pp = extract(landTypeTwo, fireLocations, cellnumbers = FALSE),
  weather = extract(weather, fireLocations, cellnumbers = FALSE)
)

# Create the dataset necessary to fit the statistical model fire escapes
#dataFireSense_EscapeFit <- tibble(
#  landtype_1_pp = extract(landTypeOne, fireLocations),
#  landtype_2_pp = extract(landTypeTwo, fireLocations),
#  weather = extract(weather, fireLocations)
#)

b1 <- .01
b2 <- -1
b3 <- 1

escapeProb <- with(dataFireSense_EscapeFit, binomial()$linkinv(weather * b1 + landtype_1_pp * b2 + landtype_2_pp * b3))
escaped <- rbinom(n = nbFires, size = 1, prob = escapeProb)

dataFireSense_EscapeFit[, `:=`(escaped = escaped, n_fires = 1)]
# dataFireSense_EscapeFit <- mutate(dataFireSense_EscapeFit, escaped = escaped, n_fires = 1)

# Create the dataset necessary to fit the statistical model of the fire size distribution
# Calculate mean landcover and weather conditions around the locations of escaped fires
buf_around_loc_escaped <- gBuffer(fireLocations[as.logical(escaped)], byid = TRUE, width = .03)

landtypeTypeOneBufMn <- extract(landTypeOne, buf_around_loc_escaped) %>% 
  lapply(mean) %>% 
  unlist

landtypeTypeTwoBufMn <- extract(landTypeTwo, buf_around_loc_escaped) %>% 
  lapply(mean) %>% 
  unlist

weatherBufMn <- extract(weather, buf_around_loc_escaped) %>%
  lapply(mean) %>%
  unlist

dataFireSense_SizeFit <- data.table(
  landtype_1_pp = landtypeTypeOneBufMn,
  landtype_2_pp = landtypeTypeTwoBufMn,
  weather = weatherBufMn
)

b1_l <- -.03
b2_l <- 4
b3_l <- -2

lambda <- with(dataFireSense_SizeFit, exp(weather * b1_l + landtype_1_pp * b2_l + landtype_2_pp * b3_l))

b1_t <- .02
b2_t <- 4
b3_t <- 2
  
theta <- with(dataFireSense_SizeFit, exp(weather * b1_t + landtype_1_pp * b2_t + landtype_2_pp * b3_t))

fireSize <- round(rtappareto(n = length(lambda), lambda = lambda, theta = theta, a = 1))

dataFireSense_SizeFit[, fire_size := fireSize]
# dataFireSense_SizeFit <- mutate(dataFireSense_SizeFit, fire_size = fireSize)

# Create the fire attribute dataset that describes the starting locations 
# and the size of the fires to be spread. This is needed to fit the statistical model of spread probabilities
fireAttributesFireSense_SpreadFit <- SpatialPointsDataFrame(fireLocations[as.logical(escaped)],
                                                            data = data.frame(size = fireSize))
```

## Fit the statistical model of fire ignitions

```{r fit_ignition_model, message=FALSE}
modules <- list("fireSense_IgnitionFit")

times <- list(start = 1, end = 1)

# Define fireSense_IgnitionFit module inputs
objects <- list(dataFireSense_IgnitionFit = dataFireSense_IgnitionFit)

# Define fireSense_IgnitionFit module parameters
formula <- formula(n_fires ~ landtype_1_pp:weather + landtype_2_pp:weather - 1)
family <- poisson(link = "identity")
ub <- list(coef = 1)
dataObjName <- "dataFireSense_IgnitionFit"
trace <- 1
iterDEoptim <- 60
iterNlminb <- 100  
cores <- 50

parameters <- list(
  fireSense_IgnitionFit = list(
    formula = formula,           # Formula of the statistical model
    family = family,             # Distribution family, here the negative binomial distribution
    ub = ub,                     # Upper bounds for coefficients to be estimated
    data = dataObjName,          # Name of the data.frame containing the variables in the statistical model. By default "dataFireSense_IgnitionFit"
    trace = trace,               # Print progress every 10 iterations
    iterDEoptim = iterDEoptim,   # Integer defining the maximum number of iterations allowed for the DEoptim optimizer
    iterNlminb = iterNlminb,     # Number of trials, or searches, to be performed by the nlminb optimizer in order to find the best solution
    cores = cores              # Number of logical cores to use for parallel computation during optimization
  )
)

# Run the simulation
sim <- simInitAndSpades(
  modules = modules,
  params = parameters,
  objects = objects,
  times = times
)

fireSense_IgnitionFitted <- sim$fireSense_IgnitionFitted # Extract the fitted model from the sim object
```

## Fit the statistical model of fire escapes

```{r fit_escape_model, message=FALSE}
modules <- list("fireSense_EscapeFit")

times <- list(start = 1, end = 1)

# Define fireSense_EscapeFit module inputs
objects <- list(dataFireSense_EscapeFit = dataFireSense_EscapeFit)

# Define fireSense_EscapeFit module parameters
formula <- formula(cbind(escaped, n_fires - escaped) ~ landtype_1_pp + landtype_2_pp + weather - 1)
dataObjName <- "dataFireSense_EscapeFit"

parameters <- list(
  fireSense_EscapeFit = list(
    formula = formula, # Formula of the statistical model
    data = dataObjName # Name of the data.frame containing the variables in the statistical model. By default "dataFireSense_EscapeFit"
  )
)

# Run the simulation
sim <- simInitAndSpades(
  modules = modules,
  params = parameters,
  objects = objects,
  times = times
)

fireSense_EscapeFitted <- sim$fireSense_EscapeFitted # Extract the fitted model from the sim object
```

## Fit the statistical model of the spread probabilities

The purpose of this step is to derive variable fire spread probabilities at the cell level, both spatially and temporally, which are sensitive to environmental drivers that control fire size.

## Fit the statistical model of the fire size distribution

Although the use of fireSense_SizeFit is optional, it can be used to reduce the dimensionality of the problem, which helps speed up and facilitate convergence when using the fireSense_SpreadFit module later.
Here, we use the fireSense_SizeFit module to model the relation between fire sizes and flammability conditions as represented by the drought index and the type of fuels.
We then use the fireSense_SpreadFit module to convert the fitted equations produced by fireSense_SizeFit into variable probabilities of fire spread at the cell level, which are also sensitive to environmental drivers that control fire size.

```{r fit_size_distribution_model, message=FALSE}
modules <- list("fireSense_SizeFit")

times <- list(start = 1, end = 1)

# Define fireSense_SizeFit module inputs
objects <- list(dataFireSense_SizeFit = dataFireSense_SizeFit)

# Define fireSense_SizeFit module parameters
formula <- formula(fire_size ~ landtype_1_pp + landtype_2_pp + weather - 1)
dataObjName <- "dataFireSense_SizeFit"

parameters <- list(
  fireSense_SizeFit = list(
    formula = list(     # Formulas of the statistical model
      beta = formula,   # The formulas for the beta and theta parameters of the tapered Pareto.
      theta = formula   # They do not have to be identical, even if it's the case here
    ),
    data = dataObjName, # Name of the data.frame containing the variables in the statistical model. By default "dataFireSense_SizeFit"
    link = list(
      beta = "log",     # Link function for the beta parameter of the tapered Pareto
      theta = "log"     # Link function for the theta parameter of the tapered Pareto
    ),
    a = 1,              # Lower truncation point a of the tapered Pareto
    ub = list(
      beta = 10,
      theta = 10
    ),
    itermax = 2000,
    trace = 100
  )
)

# Run the simulation
sim <- simInitAndSpades(
  modules = modules,
  params = parameters,
  objects = objects,
  times = times
)

fireSense_SizeFitted <- sim$fireSense_SizeFitted # Extract the fitted model from the sim object
```

### Produce the maps of the tapered Pareto parameters

```{r produce_maps_tp_params, message=FALSE}
modules <- list("fireSense_SizePredict")

times <- list(start = 1, end = 1)

# Define fireSense_SizePredict module inputs
objects <- list(
  fireSense_SizeFitted = fireSense_SizeFitted,
  landtype_1_pp = landTypeOne,
  landtype_2_pp = landTypeTwo,
  weather = weather
)

# Define fireSense_SizePredict module outputs
outputs <- rbind(
  data.frame(
    file = "fireSense_SizePredicted_Beta.tif",
    fun = "writeRaster",
    objectName = "fireSense_SizePredicted_Beta",
    package = "raster",
    saveTime = 1
  ),
  data.frame(
    file = "fireSense_SizePredicted_Theta.tif",
    fun = "writeRaster",
    objectName = "fireSense_SizePredicted_Theta",
    package = "raster",
    saveTime = 1
  )
)

# Define fireSense_SizePredict module parameters
parameters <- list(
  fireSense_SizePredict = list(
    data = c("landtype_1_pp", "landtype_2_pp", "weather"),
    modelName = "fireSense_SizeFitted" # This is the default
  )
)

# Run the simulation
sim <- simInitAndSpades(
  modules = modules,
  objects = objects,
  outputs = outputs,
  params = parameters,
  times = times
)
```

### Use predicted parameters of the tapered Pareto as inputs to the statistical model of the spread probabilities

Here, we use the `fireSense_SpreadFit` module to convert them into spatially variable spread probabilities. 

```{r fit_spread_model, message=FALSE}
modules <- list("fireSense_SpreadFit")

times <- list(start = 1, end = 1)

# Define fireSense_SpreadFit module inputs
inputs <- rbind(
  # tapered Pareto's beta
  data.frame(
    files = dir(
      getPaths()$outputPath,
      pattern = "fireSense_SizePredicted_Beta", 
      full.names = TRUE
    ),
    functions = "raster::raster",
    objectName = "beta",
    loadTime = 1
  ),
  # tapered Pareto's theta
  data.frame(
    files = dir(
      getPaths()$outputPath,
      pattern = "fireSense_SizePredicted_Theta", 
      full.names = TRUE
    ),
    functions = "raster::raster",
    objectName = "theta",
    loadTime = 1
  )
)
  
objects <- list(fireAttributesFireSense_SpreadFit = fireAttributesFireSense_SpreadFit)

# Define fireSense_SpreadFit module parameters
formula <- formula(~ I(1/beta) + log(theta) - 1)

parameters <- list(
  fireSense_SpreadFit = list(
    formula = formula, # Formula of the statistical model
    data = c("beta", "theta"),
    lower = c(.01, 0, .1, .3, .001, .001),
    upper = c(.20, .1, 10, 4., .300, .300),
    cores = 8
  )
)

# Run the simulation
sim <- simInitAndSpades(
  inputs = inputs,
  modules = modules,
  objects = objects,
  params = parameters,
  times = times
)

fireSense_SpreadFitted <- sim$fireSense_SpreadFitted # Extract the fitted model from the sim object
```

## Predict maps of fire ignition rates, escape and spread probabilities

### Fire ignition rates

```{r predict_ignition_rates, message=FALSE}
modules <- list("fireSense_IgnitionPredict")

times <- list(start = 1, end = 1)

# Define fireSense_IgnitionPredict module inputs
objects <- list(
  fireSense_IgnitionFitted = fireSense_IgnitionFitted,
  landtype_1_pp = landTypeOne,
  landtype_2_pp = landTypeTwo,
  weather = weather
)

# Define fireSense_IgnitionPredict module outputs
outputs <- rbind(
  data.frame(
    file = paste0("fireSense_IgnitionPredicted.tif"),
    fun = "writeRaster",
    objectName = "fireSense_IgnitionPredicted",
    package = "raster",
    saveTime = 1
  )
)

# Define fireSense_IgnitionPredict module parameters
parameters <- list(
  fireSense_IgnitionPredict = list(
    data = c("landtype_1_pp", "landtype_2_pp", "weather"),
    modelName = "fireSense_IgnitionFitted" # This is the default
  )
)

# Run the simulation
sim <- simInitAndSpades(
  modules = modules,
  objects = objects,
  outputs = outputs,
  params = parameters,
  times = times
)
```

### Fire escape probabilities

```{r predict_escape_probabilities, message=FALSE}
modules <- list("fireSense_EscapePredict")

times <- list(start = 1, end = 1)

# Define fireSense_EscapePredict module inputs
objects <- list(
  fireSense_EscapeFitted = fireSense_EscapeFitted,
  landtype_1_pp = landTypeOne,
  landtype_2_pp = landTypeTwo,
  weather = weather
)

# Define fireSense_EscapePredict module outputs
outputs <- rbind(
  data.frame(
    file = paste0("fireSense_EscapePredicted.tif"),
    fun = "writeRaster",
    objectName = "fireSense_EscapePredicted",
    package = "raster",
    saveTime = 1
  )
)

# Define fireSense_EscapePredict module parameters
parameters <- list(
  fireSense_EscapePredict = list(
    data = c("landtype_1_pp", "landtype_2_pp", "weather"),
    modelName = "fireSense_EscapeFitted" # This is the default
  )
)

# Run the simulation
sim <- simInitAndSpades(
  modules = modules,
  objects = objects,
  outputs = outputs,
  params = parameters,
  times = times
)
```

### Fire spread probabilities

```{r predict_spread_probabilities, message=FALSE}
modules <- list("fireSense_SpreadPredict")

times <- list(start = 1, end = 1)

# Define fireSense_SpreadPredict module inputs
# tapered Pareto's beta
inputs <- rbind(
  data.frame(
    files = dir(
      getPaths()$outputPath,
      pattern = "fireSense_SizePredicted_Beta", 
      full.names = TRUE
    ),
    functions = "raster::raster",
    objectName = "beta",
    loadTime = 1
  ),
  # tapered Pareto's theta
  data.frame(
    files = dir(
      getPaths()$outputPath,
      pattern = "fireSense_SizePredicted_Theta", 
      full.names = TRUE
    ),
    functions = "raster::raster",
    objectName = "theta",
    loadTime = 1
  )
)

objects <- list(fireSense_SpreadFitted = fireSense_SpreadFitted)

# Define fireSense_SpreadPredict module outputs
outputs <- rbind(
  data.frame(
    file = paste0("fireSense_SpreadPredicted.tif"),
    fun = "writeRaster",
    objectName = "fireSense_SpreadPredicted",
    package = "raster",
    saveTime = 1
  )
)

# Define fireSense_SpreadPredict module parameters
parameters <- list(
  fireSense_SpreadPredict = list(
    data = c("landtype_1_pp", "landtype_2_pp", "weather"),
    modelName = "fireSense_SpreadFitted" # This is the default
  )
)

# Run the simulation
sim <- simInitAndSpades(
  modules = modules,
  objects = objects,
  outputs = outputs,
  params = parameters,
  times = times
)
```

## Run fireSense, finally!

```{r run_fireSense, message=FALSE}
modules <- list("fireSense")

times <- list(start = 1, end = 1)

# Define fireSense module inputs
inputs <- rbind(
  # Fire ignition rates
  data.frame(
    files = dir(
      getPaths()$outputPath,
      pattern = "fireSense_IgnitionPredicted", 
      full.names = TRUE
    ),
    functions = "raster::raster",
    objectName = "ignitionProbRaster",
    loadTime = 1
  ),
  # Fire escape probabilities
  data.frame(
    files = dir(
      getPaths()$outputPath,
      pattern = "fireSense_EscapePredicted", 
      full.names = TRUE
    ),
    functions = "raster::raster",
    objectName = "escapeProbRaster",
    loadTime = 1
  ),
  # Fire spread probabilities
  data.frame(
    files = dir(
      getPaths()$outputPath,
      pattern = "fireSense_SpreadPredicted", 
      full.names = TRUE
    ),
    functions = "raster::raster",
    objectName = "spreadProbRaster",
    loadTime = 1
  )
)

# Run the simulation
sim <- simInitAndSpades(
  inputs = inputs,
  modules = modules,
  objects = objects,
  times = times
)

burnMap <- sim$burnMap

Plot(burnMap)
```

Note that all the steps above could be done in one simulation / SpaDES call.

# Additional resources

Read modules metadata, all module parameters are documented in `.Rmd` files.

# References

<!-- automatically generated -->
